# Panel B

```{r }
so <- tar_read(filtered_cleaned_so_6743_manual)

so_s <- ComputeGeneSetScores(
  so = so,
  assay = 'SCT',
  datatype = 'counts',
  weighting_funcs = all_weighting_funcs['unweighted'],
  genesets = map(mono_gs, ~intersect(.x, true_MRs)),
  simplify_names = T
)

if (F) {
  so_s[['sga']] <- paste(unlist(so_s[['stim_group']]), '-', 
    unlist(so_s[['dbscan_cluster']]))
}

so_s <- order_stim_group(so_s)
so_s[['sga']] <- so_s[['stim_group']]
# so_s@meta.data$stim_group
# so_s@meta.data$sga
fns <- rownames(so_s[['GS']])
fns <- c('IFNy-mono', 'TNFa-mono')

GS_p_dat <- assay2dtf(so_s, 'GS', 
  s_vars = c('condition_name', 'sample_origin', 'ifn_conc',
    'sn_dilution', 'percent.mt', 'tnf_conc', 'duration',
    'dbscan_cluster', 'sga'))
```

```{r mouse-sc-violins, eval=F, warning=FALSE, fig.width = 12/2.56, fig.height = 10/2.56}
if (interactive())
  source('~/MirjamHoekstra/R/init.R')
purrr::walk(fns, function(fn) {
  # p <- plot_vln(so_s, fn = fn, 
  #   assay = 'GS', group_var = 'sga', pt.size = .02) +
  p <- plot_vln(GS_p_dat, fn = fn, 
    group_var = 'sga', pt.size = .02) +
    scale_x_discrete(expand = c(0, 0))
    # ggpubr::stat_compare_means(
    #   label = "p.signif",
    #   method = "wilcox.test",
    #   ref.group = 'Mix PBS',
    #   label.x.npc = .5,
    #   label.y.npc = 1,
    #   colour = 'indianred3'
    # )
    # ggtitle(with(long_gene_cluster_names, gcn[which(gc == fn)]))
  pe <- rlang::expr({ print(p) })
  print_plot_eval(!!pe, width = 12, height = 10,
    filename = file.path(out_dir,
      glue::glue('violin_{fn}.pdf')))
})
```

```{r }
if (interactive() && !test_rendering())
  source('~/MirjamHoekstra/R/init.R')

GS_p_dat <- 
  assay2dtf(so_s, 'GS') %>%
  format_gs_names() %>%
  dplyr::mutate(stim_group = factor(stim_group, 
      levels = c('Mix PBS', 'Ag-GAS + T', 'Mix + T')))

for (fn in colnames(GS_p_dat)[1:2]) {
  mk_ann_GS_vln(
    p_dat = GS_p_dat,
    cn_mode = 'all',
    fn = fn,
    out_dir = out_dir,
    experiment = '6743',
    return_mode = 'file',
    p_values = 'reference',
    plot_lgd = F
  )
}
```

```{r }
compute_all_pairwise_medians(GS_p_dat) %>%
  write_tsv(file.path(out_dir, 'pairwise_gs_comp_stats_6743_mouse_in_vivo.tsv'))
```

# Panel C 

Test neighbourhoods for enrichment of a condition

```{r }
# list.files('~/MirjamHoekstra/R')
if (interactive() && !test_rendering())
  source('~/MirjamHoekstra/R/init.R')

out_dir <- file.path(Sys.getenv('img_dir'), 'fig4')
dir.create(out_dir, showWarnings = F)

mouse_genes <- tar_read(mono_reporters_mouse)$gene
so <- tar_read(filtered_cleaned_so_6743_manual)
if (F) {
  sce <- tar_read(sce_mouse)
} else {
  sce <- tar_read(sce_mouse_var_genes)
}
library(miloR)
library(SingleCellExperiment)
reducedDim(sce)

tM <- tar_read(tM_mouse)
# tar_make(names = sce_mouse_var_genes)
# limma_MRs <- tar_read(stringent_limma_MRs_mouse)
sample_annotation <- tar_read(sample_annotation_mouse)
exp5892 <- tar_read(bulk_5892_so)

if (!exists('so') || so@meta.data$exp[1] != '6743') {
  so <- tar_read(filtered_so_6743)
  sg_levs <- c('Mix PBS', 'Ag-GAS + T', 'Mix + T')
  stopifnot(all(so@meta.data$stim_group %in% sg_levs))
  so@meta.data$stim_group <- factor(so@meta.data$stim_group,
    levels = sg_levs)
}

## Remove IFNa
limma_MR_params <- tar_read(limma_MR_params_mouse)
M_cpm <- tar_read(M_cpm_mouse)
tM <- tar_read(tM_mouse)[, -2]
# tM <- tar_read(tM_mouse)
# tM <- tar_read(tM_mouse)[, c(3, 5, 6, 7)]
RC  <- tM >= limma_MR_params$LL
# RC  <- tM >= 3
NRC <- tM < limma_MR_params$UL
MRs <- names(which(
  apply(RC[, 1:5], 1, sum) == 1 &
  apply(NRC[, 1:5], 1, sum) >= 4 &
  # abs(apply(tM[, 4:5], 1, sum) - tM[, 6]) <= limma_MR_params$SD &
  T
))

# non_combo_MRs <- names(which(
#   apply(RC[, 1:5], 1, sum) == 1 &
#   apply(NRC[, 1:5], 1, sum) >= 4
# ))

f_genes <- rownames(tar_read(f_gene_var_mouse))
gene_sets <- 
  MRs %>%
  # non_combo_MRs %>%
  { tapply(., colnames(tM)[apply(RC[., 1:5], 1, which.max)], list) } %>%
  map(~intersect(.x, f_genes)) %>%
  { setNames(., gsub('.* ng/ml ', '', names(.))) }

gene_sets[['IFNy']] <- mirjam_genes_220516$IFNy_mono
gene_sets[['TNFa']] <- mirjam_genes_220516$TNFa_mono

map(gene_sets, length)
sum(map_int(gene_sets, length))

if (interactive() && !test_rendering())
  source('~/MirjamHoekstra/R/init.R')
```

```{r}
library(miloR)
library(SummarizedExperiment)

print_plot_eval(print(plotNhoodSizeHist(sce)),
  width = 17.4, height = 10,
  filename = file.path(out_dir,
    glue::glue('mouse_Nhood_sizes.pdf')))

sce <- countCells(
  sce,
  meta.data = data.frame(colData(sce)),
  samples = c('condition_i')
  # samples = c('stim_group')
)
# head(nhoodCounts(sce))

design_M <-
  data.frame(colData(sce)) %>%
  # tibble::rownames_to_column('sample_name') %>%
  dplyr::select(
    condition_i, duration, stim_group,
    matches('mouse')
  ) %>%
  dplyr::distinct() %>%
  dplyr::arrange(condition_i) %>%
  set_rownames(NULL) %>%
  # { set_rownames(., .$sample_name) }
  { . }
design_M <- model.matrix(~1 + stim_group, design_M) %>%
  as.data.frame()
colnames(design_M) <- colnames(design_M) %>%
  stringr::str_replace_all(' |\\+|-', '_')

all_NH_test_DA <-
  colnames(design_M)[2:ncol(design_M)] %>%
  auto_name() %>%
  map(function(cn) {
    testNhoods(
      sce,
      model.contrasts = cn,
      # model.contrasts = 'stim_groupAg_GAS___T',
      design = ~ stim_groupMix___T + stim_groupAg_GAS___T,
      design.df = as.data.frame(design_M)
    ) %>%
    dplyr::mutate(Nhood = unlist(nhoodIndex(sce))) %>%
    { set_rownames(., .$Nhood) } %>%
    # dplyr::arrange(desc(logFC)) %>%
    { . }
  })

sce_s <- sce[, colData(sce)$stim_group == 'Mix + T']
sce@nhoods[colData(sce)$stim_group == 'Mix + T', ]

NH_test_DA <- all_NH_test_DA[['stim_groupMix___T']]
# all_NH_test_DA[[2]]$SpatialFDR <= 0.05
```

```{r, warning=FALSE}
# nrow(all_NH_test_DA[[vn]])
if (interactive() && !test_rendering())
  source('~/MirjamHoekstra/R/init.R')

for (vn in names(all_NH_test_DA)) {
  mcat <- function(...) cat(..., '\n')

  mcat('\nStim group:', vn)
  mcat('\nEnriched')
  idx <- 
    which(with(all_NH_test_DA[[vn]], SpatialFDR <= .05 & logFC > 0))
  mcat(length(idx), 'Nhood(s)')
  mcat(frac_included_cells(sce, 
      allowed_Nhoods = as.character(all_NH_test_DA[[vn]]$Nhood[idx])))
  mcat(frac_included_cells(sce, 
      allowed_cells = which(colData(sce)$stim_group == 'Mix + T'), 
      allowed_Nhoods = as.character(all_NH_test_DA[[vn]]$Nhood[idx])))

  mcat('\nDepleted')
  idx <- 
    which(with(all_NH_test_DA[[vn]], SpatialFDR <= .05 & logFC < 0))
  mcat(length(idx), 'Nhood(s)')
  mcat(frac_included_cells(sce, 
      allowed_Nhoods = as.character(all_NH_test_DA[[vn]]$Nhood[idx])))
  mcat(frac_included_cells(sce, 
      allowed_cells = which(colData(sce)$stim_group == 'Mix + T'), 
      allowed_Nhoods = as.character(all_NH_test_DA[[vn]]$Nhood[idx])))
  mcat(frac_included_cells(sce, 
      allowed_cells = which(colData(sce)$stim_group == 'Mix PBS'), 
      allowed_Nhoods = as.character(all_NH_test_DA[[vn]]$Nhood[idx])))
  mcat(frac_included_cells(sce, 
      allowed_cells = which(colData(sce)$stim_group == 'Ag-GAS + T'), 
      allowed_Nhoods = as.character(all_NH_test_DA[[vn]]$Nhood[idx])))
}
# NH_test_DA <-
#   testNhoods(
#     sce,
#     model.contrasts = 'stim_groupMix___T',
#     # model.contrasts = 'stim_groupAg_GAS___T',
#     design = ~ stim_groupMix___T + stim_groupAg_GAS___T,
#     design.df = as.data.frame(design_M)
#   ) %>%
#   dplyr::arrange(desc(logFC))
```


```{r }
if (interactive() && !test_rendering())
  source('~/MirjamHoekstra/R/init.R')
library(miloR)

rename_mouse_stimgroup <- function(dtf) {
  dtf <- order_stim_group(dtf)
  dtf$stim_group <- forcats::lvls_revalue(
    dtf$stim_group,
    new_levels = c('Control', 'Ag- T-cell', 'Ag+ T-cell')
  )
  return(dtf)
}

p_dat <-
  as.data.frame(cbind(colData(sce), reducedDim(sce, 'UMAP'))) %>%
  order_stim_group() %>%
  order_duration() %>% 
  rename_mouse_stimgroup() %>%
  { . }
if (F) {
  p_dat$UMAP_1
  p_dat$UMAP_2
  p_dat$stim_group
  p_dat$duration
}

p <-
  plot_dim_reduc(
    p_dat,
    coord_regex = 'UMAP_',
    colour_var = 'stim_group',
    use_stim_group_cols = FALSE,
    shape_var = NULL
  ) +
  theme(legend.position = 'bottom') +
  guides(
    color = guide_legend(
      ncol = 1,
      override.aes = list(alpha = 1)
    )
  ) +
  theme()

if (interactive() && !test_rendering())
  source('~/MirjamHoekstra/R/init.R')
# p2 <- plotNhoodGraphDA(sce, da_results, layout="umap",alpha=0.05)
p2 <- plotNhoodGraphDA(
  x = sce,
  layout = 'UMAP',
  # milo_res = NH_test_DA[match(unlist(nhoodIndex(sce)),
  #   NH_test_DA$Nhood), ],
  milo_res = all_NH_test_DA[[1]],
  alpha = .05) +
  theme_cyto_inf() +
  theme(plot.title = element_text(hjust = .5)) +
  scale_fill_gradient2(
    name = 'logFC DA', 
    limits = range(all_NH_test_DA[[2]]$logFC)
  ) +
  theme_tabula_rasa +
  ggtitle('Ag- T-cell')

p3 <- plotNhoodGraphDA(
  x = sce,
  layout = 'UMAP',
  # milo_res = NH_test_DA[match(unlist(nhoodIndex(sce)),
  #   NH_test_DA$Nhood), ],
  milo_res = all_NH_test_DA[[2]],
  alpha = .05) +
  theme_cyto_inf() +
  theme_tabula_rasa +
  theme(plot.title = element_text(hjust = .5)) +
  scale_fill_gradient2(
    name = 'logFC DA', 
    limits = range(all_NH_test_DA[[2]]$logFC)
  ) +
  ggtitle('Ag+ T-cell')

print_plot_eval(
  print((p + plot_spacer()) / (p2 + p3 + plot_layout(guides = 'collect'))),
  width = 17.4, height = 20,
  filename = file.path(
    out_dir,
    glue::glue('mouse_Nhood_HL_UMAP.pdf')
  )
)
```


```{r, echo = FALSE options}
gene = 'Ckb'
gene = 'Itga5'
gene = 'Smad7'
gene = 'Gjb3'
colData(sce)[[gene]] <- NA_real_
colData(sce)[[gene]][as.integer(colnames(nhoodExpression(sce)))] <-
  # unname(nhoodExpression(sce)[gene, ])
  unname(min_max_scaling(nhoodExpression(sce)[gene, ]))

colData(sce)[[gene]]

p0 <- plotNhoodGraph(
  x = sce,
  colour_by = gene) +
  theme_cyto_inf() +
  gg_tabula_rasa +
  xlab('') +
  ylab('') +
  guides(size = 'none', edge = 'none') +
  theme(legend.position = 'right', legend.direction = 'vertical')

print_plot_eval(print(p0),
  width = 17.4, height = 10,
  filename = file.path(out_dir,
    glue::glue('{gene}_Nhoods.pdf')))
```

# Panel D

Compare T-cell exposed to non-exposed using Milo
Globally, which genes associate with T-cell exposed cells?

```{r }
# gene_idxs <- which(dge_smp$logCPM_6 >= mean(dge_smp$logCPM_6))

Nhood_expression <- calcNhoodExpression(sce)

M_cpm <- tar_read(M_cpm_mouse)
# meanE <- apply(M_cpm, 1, mean)
# maxE <- apply(M_cpm, 1, max)
# gene_idxs <- which(meanE >= mean(maxE))
# gene_idxs <- which(maxE >= mean(maxE))
# setdiff(names(cors), names(gene_idxs))
# setdiff(names(gene_idxs), names(cors))

cors_fn <- file.path(rds_dir, 'DA_gene_cors_mouse.rds')
cors_fn <- file.path(rds_dir, 'DA_gene_cors_mouse_all_genes.rds')
cors_fn <- file.path(rds_dir, 'DA_gene_cors_mouse_all_genes_var_genes.rds')
if (!file.exists(cors_fn)) {
  cors <-
    rownames(assay(Nhood_expression, 'logcounts')) %>%
    intersect(rownames(M_cpm)) %>%
    auto_name() %>%
    purrr::map(function(g) {
      cor(
        NH_test_DA$logFC,
        assay(Nhood_expression, 'logcounts')[g, NH_test_DA$Nhood],
        method = 'spearman'
      )
    }) %>%
    # setNames(rownames(Nhood_expression)[gene_idxs])
    { . }
  cors <- unlist(cors)
  # quantile(cors, seq(0, 1, .05))
  saveRDS(cors, cors_fn)
} else {
  cors <- readRDS(cors_fn)
}
```

# Panel B

```{r, warning=FALSE}
source('~/MirjamHoekstra/R/plot_DA_panel.R')
N = 250
feats <-
  names(sort(abs(cors))) %>%
  { intersect(., rownames(Nhood_expression)) } %>%
  tail(n = N)
plot_DA_panel(
  genes = feats,
  row_title = glue::glue('{N} most strongly correlated genes'),
  of = file.path(out_dir,
    glue::glue('top{N}_DA_panel.pdf'))
)

cors[c('Vim', 'Fn1', 'Cdh2')]
cors[c('Cdh1')]
summary(cors[hallmark_gs[["Interferon gamma response"]]])
summary(cors[hallmark_gs[["TNFa signaling via NFkB"]]])
summary(cors[hallmark_gs[["TGF beta signaling"]]])
summary(cors[hallmark_gs[["Epithelial mesenchymal transition"]]])
summary(cors)


summary(cors[intersect(names(cors), gene_sets[['TGFb']])])
table(sign(cors[intersect(names(cors), gene_sets[['TGFb']])]))

# summary(cors[intersect(names(cors), TGFb_MRs)])
# table(sign(cors[intersect(names(cors), TGFb_MRs)]))

plot_DA_panel(
  genes = gene_sets[['TGFb']],
  row_title = 'TGFb mono-responsive genes',
  of = file.path(out_dir,
    glue::glue('TGFb_responsive_genes_DA_panel.pdf'))
)
```

Not in paper: pathway analysis on correlated genes

```{r }
devtools::load_all(file.path('~/libs', 'maartenutils'))
library(genesets)

hallmark_gs <-
  filter_gmt(
    pattern = 'HALLMARK',
    # gmt_pattern = 'msigdb',
    gmt_pattern = 'mh.all',
    type = 'symbols'
  )
caplist <- caplist_def %>%
  c('via')
names(hallmark_gs) <- names(hallmark_gs) %>%
  stringr::str_replace('HALLMARK_', '') %>%
  tolower() %>%
  # stringr::str_replace('_', ' ') %>%
  simple_cap(cap_first_word_only = T, caplist = caplist)
names(hallmark_gs)
```

```{r }
if (interactive() && !test_rendering())
  source('~/MirjamHoekstra/R/init.R')
source('~/MirjamHoekstra/R/gsea.R')

head(sort(cors), n = 30)

p <- gsea_table(
  cors[!is.na(cors)],
  analysis_title = glue::glue('T-cell pressure correlation'),
  min_p = .05,
  max_pathways = 50,
  colwidths = c(3, 3, 0.8, 0, 1.2),
  return_grob = T,
  collapse = FALSE,
  pathways = hallmark_gs
)

print_plot_eval(grid.draw(p),
  width = 17.4, height = 10,
  filename = file.path(out_dir,
    glue::glue('T_cell_pressure_cor_Hallmark.pdf')))
```

## Look within experimental condition rather than across them

Not in paper. Scatter plots of gene set scores. Do we also see a difference in
TGFb/IFNy when stratifying purely on
condition?

```{r }
colnames(M_cpm) <- sample_annotation$sample_name
``` 

```{r }
MAGIC <- readr::read_csv(file.path(data_dir, '6743_Mix_T_MAGIC.csv'))
MAGIC <- MAGIC[, -1] %>%
  { .[, intersect(colnames(.), unlist(gene_sets))] }

gs_p_dat <-
  t(ComputeGeneSetScores(
    M = t(MAGIC), 
    genesets = gene_sets,
    weighting_funcs = all_weighting_funcs[c('unweighted')]
  )) %>%
  as.data.frame() %>%
  { . }

p1 <- ggplot(gs_p_dat, 
  aes_string(y = 'TGFb', x = 'IFNy', colour = 'TNFa')) + 
  geom_point(alpha = .3, size = .5) +
  geom_smooth(method = 'lm') +
  scale_colour_viridis_c()
print_plot_eval(print(p1),
  width = 12, height = 10,
  filename = file.path(out_dir,
    glue::glue('gs_scores_MAGIC.pdf')))

p1 <- ggplot(gs_p_dat, 
  aes_string(y = 'TGFb', x = 'IFNy', colour = 'TNFa')) + 
  geom_point(alpha = .3, size = .5) +
  geom_smooth(method = 'lm') +
  scale_colour_viridis_c()
print_plot_eval(print(p1),
  width = 12, height = 10,
  filename = file.path(out_dir,
    glue::glue('gs_scores_MAGIC.pdf')))
```

```{r, warning=FALSE}
geneset_labels <- 
  tibble::enframe(gene_sets, 'gs', 'gene') %>%
  dplyr::mutate(gs = factor(gs, 
      levels = c('TNFa', 'IFNy', 'TGFb'))) %>%
  tidyr::unnest(gene)

plot_corM <- function(corM, fn = file.path(out_dir,
      glue::glue('exp6743_T_cell_MAGIC_corM.pdf'))) {
  HM <- Heatmap(
    corM,
    # column_title = cn,
    name = 'Correlation',
    col <- circlize::colorRamp2(c(-1, 0, 1), 
      c('navyblue', 'white', 'indianred3')),
    left_annotation = rowAnnotation(
      'Gene set' = anno_simple(as.character(geneset_labels$gs[
        match(rownames(corM), geneset_labels$gene)]),
      col = c('TNFa' = 'blue', 'IFNy' = 'red', 'TGFb' = 'yellow'),
      gp = gpar(fontsize = 6))
    ),
    top_annotation = columnAnnotation(
      'Gene set' = anno_simple(as.character(geneset_labels$gs[
        match(rownames(corM), geneset_labels$gene)]),
      col = c('TNFa' = 'blue', 'IFNy' = 'red', 'TGFb' = 'yellow'),
      gp = gpar(fontsize = 6))
    ),
  )
  print_plot_eval(
    draw(HM, merge_legend = TRUE, heatmap_legend_side = 'top'),
    # width = 17.4, 
    width = 17.4, 
    height = 15,
    filename = fn
  )
}

plot_corM(cor(MAGIC, method = 'pearson'), 
  file.path(out_dir, glue::glue('exp6743_T_cell_corM_MAGIC.pdf'))
)

M %>%
  { .[, intersect(colnames(.), unlist(gene_sets))] } %>%
  cor(method = 'pearson') %>%
  plot_corM(file.path(out_dir, glue::glue('exp6743_T_cell_corM.pdf')))
```

```{r }
p_dat <- as.data.frame(MAGIC) %>%
  add_umap()

g1 = 'Galnt1'
g2 = 'Fgl2'
g3 = 'Cx3cl1'
colnames(p_dat)

g1 = 'Tnnt2'
g2 = 'Loxl3'
g2 = 'Bambi'
g2 = 'Batf2'
g2 = 'Clic5'
g2 = 'Cx3cl1'
g3 = 'Col12a1'
g3 = 'Galnt1'
p1 <- ggplot(p_dat, 
  aes_string(x = g1, y = g2, colour = g3)) + 
  geom_point(alpha = .3, size = .5) +
  scale_colour_viridis_c()
p2 <- ggplot(p_dat, 
  aes_string(x = g1, y = g3, colour = g3)) + 
  geom_point(alpha = .3, size = .5) +
  scale_colour_viridis_c()
p3 <- ggplot(p_dat, 
  aes_string(x = 'UMAP1', y = 'UMAP2', colour = g1)) + 
  gg_tabula_rasa +
  geom_point(alpha = .3, size = .5) +
  scale_colour_viridis_c()
p4 <- ggplot(p_dat, 
  aes_string(x = 'UMAP1', y = 'UMAP2', colour = g2)) + 
  gg_tabula_rasa +
  geom_point(alpha = .3, size = .5) +
  scale_colour_viridis_c()
p5 <- ggplot(p_dat, 
  aes_string(x = 'UMAP1', y = 'UMAP2', colour = g3)) + 
  gg_tabula_rasa +
  geom_point(alpha = .3, size = .5) +
  scale_colour_viridis_c()
print_plot_eval(print((p3 + p4 + p5)/ (p1 + p2)),
  width = 12, height = 10,
  filename = file.path(out_dir,
    glue::glue('{g1}-{g2}-{g3}.pdf')))
```

```{r }
p_dat <- as.data.frame(M) %>%
  add_umap()

g1 = 'Galnt1'
g2 = 'Fgl2'
g3 = 'Cx3cl1'

g1 = 'Tnnt2'
g2 = 'Loxl3'
g2 = 'Bambi'
g2 = 'Batf2'
g2 = 'Cx3cl1'
g3 = 'Col12a1'
g3 = 'Galnt1'
p1 <- ggplot(p_dat,
  aes_string(x = g1, y = g2, colour = g3)) + 
  geom_point(alpha = .3, size = .5) +
  scale_colour_viridis_c()
p3 <- ggplot(p_dat, 
  aes_string(x = 'UMAP1', y = 'UMAP2', colour = g1)) + 
  gg_tabula_rasa +
  geom_point(alpha = .3, size = .5) +
  scale_colour_viridis_c()
p4 <- ggplot(p_dat, 
  aes_string(x = 'UMAP1', y = 'UMAP2', colour = g2)) + 
  gg_tabula_rasa +
  geom_point(alpha = .3, size = .5) +
  scale_colour_viridis_c()
p2 <- ggplot(p_dat, 
  aes_string(x = 'UMAP1', y = 'UMAP2', colour = g3)) + 
  gg_tabula_rasa +
  geom_point(alpha = .3, size = .5) +
  scale_colour_viridis_c()
print_plot_eval(print((p3 + p4 + p2)/ p1),
  width = 12, height = 10,
  filename = file.path(out_dir,
    glue::glue('orig_{g1}-{g2}-{g3}.pdf')))
```

# Panel F

## Data loading and wrangling

```{r load_genesets, cache=TRUE}
if (interactive() && !test_rendering())
  source('~/MirjamHoekstra/R/init.R')


s_cols <- gen_cyto_inf_col_scale()
pal <- c(
  'none' = unname(s_cols['Unstimulated']),
  # 'TGFb' = unname(s_cols['10 ng/ml TGFb']),
  # 'TGFb' = unname(darken(s_cols['10 ng/ml TGFb'], factor=1.3)),
  'TGFb' = 'blue',
  'TGFb_not_clean' = unname(darken(s_cols['10 ng/ml TGFb'], factor=3)),
  'IFNy' = unname(s_cols['100 ng/ml IFNy'])
)

mouse_genes <- tar_read(mono_reporters_mouse)$gene
so <- tar_read(filtered_cleaned_so_6743_manual)
if (F) {
  sce <- tar_read(sce_mouse)
} else {
  sce <- tar_read(sce_mouse_var_genes)
}
library(miloR)
library(SingleCellExperiment)
# reducedDim(sce)

Nhood_expression <- calcNhoodExpression(sce)
## From fig6
cors_fn <- file.path(rds_dir, 'DA_gene_cors_mouse_all_genes_var_genes.rds')
cors <- readRDS(cors_fn)

tM <- tar_read(tM_mouse)
# tar_make(names = sce_mouse_var_genes)
# limma_MRs <- tar_read(stringent_limma_MRs_mouse)
sample_annotation <- tar_read(sample_annotation_mouse)
# sample_annotation[, c('sample_name', 'stim_group')]
exp5892 <- tar_read(bulk_5892_so)

if (!exists('so') || so@meta.data$exp[1] != '6743') {
  so <- tar_read(filtered_so_6743)
  sg_levs <- c('Mix PBS', 'Ag-GAS + T', 'Mix + T')
  stopifnot(all(so@meta.data$stim_group %in% sg_levs))
  so@meta.data$stim_group <- factor(so@meta.data$stim_group,
    levels = sg_levs)
}

## Remove IFNa
limma_MR_params <- tar_read(limma_MR_params_mouse)
M_cpm_5892 <- tar_read(M_cpm_mouse)
tM <- tar_read(tM_mouse)[, -2]
# tM <- tar_read(tM_mouse)
# tM <- tar_read(tM_mouse)[, c(3, 5, 6, 7)]
RC  <- tM >= limma_MR_params$LL
NRC <- tM < limma_MR_params$UL
MRs <- names(which(
  apply(RC[, 1:5], 1, sum) == 1 &
  apply(NRC[, 1:5], 1, sum) >= 4 &
  # abs(apply(tM[, 4:5], 1, sum) - tM[, 6]) <= limma_MR_params$SD &
  T
))

# non_combo_MRs <- names(which(
#   apply(RC[, 1:5], 1, sum) == 1 &
#   apply(NRC[, 1:5], 1, sum) >= 4
# ))

f_genes <- rownames(tar_read(f_gene_var_mouse))
gene_sets <-
  MRs %>%
  # non_combo_MRs %>%
  { tapply(., colnames(tM)[apply(RC[., 1:5], 1, which.max)], list) } %>%
  map(~intersect(.x, f_genes)) %>%
  { setNames(., gsub('.* ng/ml ', '', names(.))) }

gene_sets[['IFNy']] <- mirjam_genes_220516$IFNy_mono
gene_sets[['TNFa']] <- mirjam_genes_220516$TNFa_mono

map(gene_sets, length)
sum(map_int(gene_sets, length))

if (interactive() && !test_rendering())
  source('~/MirjamHoekstra/R/init.R')
```


```{r, warning=FALSE}
source('~/MirjamHoekstra/R/init.R')

l_img_dir <- file.path(img_dir, 'exp7296')
# dir.create(l_img_dir)

sfn <- file.path(data_dir, 'raw_exp_7296', 'genecounts.txt')
expM_7296 <-
  readr::read_tsv(sfn) %>%
  rename_with(~tolower(.x)) %>%
  rename_with(~gsub('\\d{1,2}_(.*)', '\\1', .x)) %>%
  rename_with(~gsub('ifng', 'ifny', .x)) %>%
  dplyr::select(-ensembl_gene_id, -start_position, -end_position,
    -gene_biotype, -chromosome_name, -description) %>%
  dplyr::rename(gene = external_gene_id)

# colSums(M)
# sort(M['Casp8', ])
# sort(M['Cxcl10', ])
```

```{r, include = F, warning=FALSE}
num_cols <- setdiff(colnames(expM_7296), 'gene')
duration1 <- as.numeric(gsub('(\\d{1,2})h_.*$', '\\1', num_cols))
duration2 <- as.numeric(gsub('.*_(\\d{1,2})h_.*$', '\\1', num_cols))
stim1 <- factor(gsub('\\d{1,2}h_([^_]*)_.*$', '\\1', num_cols))
stim2 <- factor(gsub('.*_([^_]*)$', '\\1', num_cols))
repl_sg <- function(v) {
  v[v=='ifny'] <- '100 ng/ml IFNy'
  v[v=='tgfb'] <- '10 ng/ml TGFb'
  v[v=='medium'] <- ''
  v[v=='mix'] <- '100 ng/ml IFNy 10 ng/ml TGFb'
  return(v)
}
sample_annotation_7296 <-
  data.frame(
    'sample_name' = setdiff(colnames(expM_7296), 'gene'),
    'duration' = duration1,
    'duration2' = duration2,
    'm_sg' = repl_sg(as.character(stim1)),
    'm_sg2' = repl_sg(as.character(stim2)),
    sample_origin = include_na_lev(v='in_vitro', el='in_vivo'),
    sample_type = include_na_lev('bulk', el='sc')
  ) %>%
  dplyr::arrange(duration, m_sg, m_sg2) %>%
  { . }

ru <- function(x) ifelse(x == '', 'Unstimulated', x)
sample_annotation_7296$condition_name <- with(sample_annotation_7296, paste0(ru(m_sg), ' ', duration, 'h - ', ru(m_sg2), ' ', duration2, 'h'))

if (FALSE) {
  writeLines(
    paste(sample_annotation_7296$condition_name, collapse=', '), 
    con=file.path(img_dir, 'conditions.txt')
  )
}
```

```{r}
sample_annotation_7296 <-
  sample_annotation_7296 %>%
  dplyr::arrange(m_sg2, duration, m_sg)

expM_7296 <-
  expM_7296[, match(c('gene', sample_annotation_7296$sample_name), colnames(expM_7296))]

M_7296 <-
  as.matrix(dplyr::select(expM_7296, -gene)) %>%
  # { .[, match(sample_annotation_7296$sample_name, colnames(.))] } %>%
  { set_colnames(., tolower(colnames(.))) } %>%
  # { . }
  set_rownames(expM_7296$gene) %>%
  edgeR::DGEList() %>%
  edgeR::calcNormFactors(method = 'TMM')

M_cpm_7296 <- edgeR::cpm(M_7296)
M_tmm_7296 <- as.matrix(RunTMM(M_7296))
stopifnot(colnames(M_cpm_7296) == sample_annotation_7296$sample_name)
if (T) {
  ridx <- colnames(M_cpm_7296) != '6h_medium_24h_medium'
  M_cpm_7296 <- M_cpm_7296[, ridx]
  M_tmm_7296 <- M_tmm_7296[, ridx]
  sample_annotation_7296 <- sample_annotation_7296[ridx, ]
}
```

```{r}
# Create a version of the data that is corrected for exposure duration. Should
# have used SummarizedExperiment or something similar here, would have saved me
# from some headaches.
M_corr_7296 <- M_tmm_7296

control_idxs <- which(
  sample_annotation_7296$m_sg == '' &
  sample_annotation_7296$m_sg2 == ''
)
# sample_annotation_7296[control_idxs, ]

unique_durations <-
  sample_annotation_7296[control_idxs, c('duration', 'duration2')]

# sample_annotation_7296 %>%
#   dplyr::distinct(duration, duration2) %>%
#   dplyr::select(duration, duration2)

deduct <- function(x, y) log2(x+.01) - log2(y+.01)

all_idxs <- c()
for (i in 1:nrow(unique_durations)) {
  idxs <- which(
    !(sample_annotation_7296$m_sg == '' & sample_annotation_7296$m_sg2 == '') &
    sample_annotation_7296$duration == unique_durations[i, 'duration'] &
    sample_annotation_7296$duration2 == unique_durations[i, 'duration2'])
  all_idxs <- c(all_idxs, idxs)
  M_corr_7296[, idxs] <-
    deduct(M_tmm_7296[, idxs], M_tmm_7296[, control_idxs[i]])
}

sa_corr_7296 <- sample_annotation_7296[all_idxs, ]

# sa_corr_7296[, c('sample_name', 'm_sg', 'm_sg2')]
M_corr_7296 <- M_corr_7296[, all_idxs]
dim(sa_corr_7296); dim(M_corr_7296)

# intersect(sa_corr_7296$sample_name, colnames(M_corr_7296))
# setdiff(sa_corr_7296$sample_name, colnames(M_corr_7296))
# setdiff(colnames(M_corr_7296), sa_corr_7296$sample_name)

sa_corr_7296 <- sa_corr_7296 %>% arrange(m_sg2, m_sg, duration)
M_corr_7296 <- M_corr_7296[, sa_corr_7296$sample_name]
stopifnot(sa_corr_7296$sample_name == colnames(M_corr_7296))

sa_corr_7296[, c('sample_name', 'm_sg', 'm_sg2')]
```

```{r}
M_5892 <- tar_read(M_mouse)
sa_5892 <- tar_read(sample_annotation_mouse)
sa_5892$sample_name <- tolower(sa_5892$sample_name)

if (T) {
  idxs <- which(!stringr::str_detect(
      as.character(sa_5892$stim_group),
      'IFNA|IL-2|A1 B2'))
  length(idxs)
  sa_5892 <- sa_5892[idxs, ]
  M_5892 <- M_5892[, idxs]
}
# sa_5892$m_sg <- 'Unstimulated in vitro'
# sa_5892$m_sg <- ''
# sa_5892$m_sg2 <- sa_5892$stim_group
# sa_5892$duration2 <- sa_5892$duration
# sa_5892$duration <- 0L

sa_5892 <- as.data.frame(order_stim_group(sa_5892))
correct_conc <- function(v, pairs=list(
  c('0.1 ng/ml TNFa', '1 ng/ml TNFa'),
  c('0.1 ng/ml IFNy', '1 ng/ml IFNy'))) {
  for (p in pairs) {
    idxs <- match(levels(v), p)
    if (any(idxs, na.rm = TRUE)) {
      detected_order <- setdiff(idxs, NA)
      if (detected_order != seq_along(p)) {
        levels(v)[which(!is.na(idxs))] <- p
      }
    }
  }
  return(v)
}
# sa_5892$m_sg2 <- correct_conc(sa_5892$m_sg2)
# sa_5892$stim_group <- correct_conc(sa_5892$stim_group)


sa_5892$m_sg <- sa_5892$stim_group
sa_5892$m_sg2 <- sa_5892$stim_group
sa_5892$duration2 <- sa_5892$duration
# sa_5892$m_sg[grepl('_1_ng_ml_ifng', sa_5892$sample_name)] <- '1 ng/ml IFNy'
# sa_5892$m_sg[grepl('_0_01_ng_ml_ifng', sa_5892$sample_name)] <- '0.01 ng/ml IFNy'
sa_5892[, c('sample_name', 'm_sg')]

M_tmm_5892 <- as.matrix(RunTMM(M_5892))
dim(M_tmm_5892)
# tail(sort(colSums(M_tmm_5892)))
```


```{r}
levels(sa_5892$m_sg2) <- 
c('Unstimulated in vitro', '6000 U/ml IL-2', '5000 U/ml IFNA1',
'10 ng/ml TGFb', '2 ng/ml lymphotoxin A1 B2', '0.1 ng/ml TNFa',
'1 ng/ml TNFa', '10 ng/ml TNFa', '0.01 ng/ml IFNy', '1 ng/ml IFNy', 
'100 ng/ml IFNy', '100 ng/ml IFNy 10 ng/ml TNFa')


sa_5892$sample_name <- tolower(sa_5892$sample_name)
sa_5892 <- dplyr::arrange(sa_5892, m_sg2, duration)
sa_5892$sample_name
M_tmm_5892 <- M_tmm_5892[, tolower(sa_5892$sample_name)]
## Check whether these values are increasing
grep('ifng', colnames(M_tmm_5892), value=T)
```

```{r}
# Create a version of the data that is corrected for exposure
# duration
M_corr_5892 <- M_tmm_5892
dim(M_corr_5892)

control_idxs <- which(
  sa_5892$m_sg2 == 'Unstimulated in vitro'
)

unique_durations <-
  sa_5892[control_idxs, c('duration', 'duration2')]

deduct <- function(x, y) log2(x+.01) - log2(y+.01)

all_idxs <- c()
for (i in 1:nrow(unique_durations)) {
  idxs <- which(
    sa_5892$m_sg2 != 'Unstimulated in vitro' &
    sa_5892$duration == unique_durations[i, 'duration'] &
    sa_5892$duration2 == unique_durations[i, 'duration2'])
  all_idxs <- c(all_idxs, idxs)
  M_corr_5892[, idxs] <-
    deduct(M_tmm_5892[, idxs], M_tmm_5892[, control_idxs[i]])
}
length(all_idxs)

sa_corr_5892 <- sa_5892[all_idxs, ]
M_corr_5892 <- M_corr_5892[, tolower(sa_corr_5892$sample_name)]


intersect(sa_corr_5892$sample_name, colnames(M_corr_5892))
setdiff(sa_corr_5892$sample_name, colnames(M_corr_5892))
setdiff(colnames(M_corr_5892), sa_corr_5892$sample_name)

stopifnot(sa_corr_5892$sample_name == colnames(M_corr_5892))

# ## Select down to the included samples in M_corr
# sa_corr_5892 <- sa_corr_5892[match(colnames(M_corr_5892), sa_corr_5892$sample_name), ]

sa_corr_5892 <- dplyr::arrange(sa_corr_5892, m_sg2, duration2, m_sg)
M_corr_5892 <- M_corr_5892[, sa_corr_5892$sample_name]
stopifnot(sa_corr_5892$sample_name == colnames(M_corr_5892))
sa_corr_5892[, c('sample_name', 'm_sg', 'm_sg2')]
sa_corr_5892$m_sg2 <- sa_corr_5892$m_sg
```

## Test reconstruction error

How much do we need of each profile to
reconstruct the Nhood expression profiles? Two metrics: the absolute
size of weights, the reconstruction error

```{r}
bulk_experiment = '5892'
bulk_experiment = '7296'
if (bulk_experiment == '7296') {
  bulk_M <- M_tmm_7296
} else if (bulk_experiment == '5892'){
  bulk_M <- M_tmm_5892
}
shared_genes <- intersect(
  rownames(nhoodExpression(Nhood_expression)),
  rownames(bulk_M)
)

NhoodE <- subset_feats(
  nhoodExpression(Nhood_expression),
  shared_genes)
stopifnot(rownames(NhoodE) == shared_genes)
if (T) {
  NhoodE <- as.matrix(RunTMM(exp(NhoodE) - 1))
} else {
  dn <- dimnames(NhoodE)
  NhoodE <- NhoodE %*% diag(10^6/colSums(NhoodE))
  dimnames(NhoodE) <- dn
}

sort(colSums(NhoodE)); tail(rowSums(NhoodE))
bulkE <- subset_feats(bulk_M, shared_genes)

if (F) {
  NhoodE <- t(scale(t(NhoodE)))
  bulkE <- t(scale(t(bulkE)))
}

if (F) {
  NhoodE <- log2(NhoodE+1)
  bulkE <- log2(bulkE+1)
}

if (T) {
  NhoodE <- log10(NhoodE+1)
  bulkE <- log10(bulkE+1)
}


stopifnot(rownames(bulkE) == shared_genes)
if (T) {
} else {
  dn <- dimnames(bulkE)
  bulkE <- bulkE %*% diag(10^6/colSums(bulkE))
  stopifnot(all(abs(colSums(bulkE) - 1e6) < 1e-5))
  dimnames(bulkE) <- dn
}

rm_genes <- union(
  names(which(apply(is.na(bulkE), 1, any))),
  names(which(apply(is.na(NhoodE), 1, any))))
shared_genes <- setdiff(shared_genes, rm_genes)

bulkE <- subset_feats(bulkE, shared_genes)
NhoodE <- subset_feats(NhoodE, shared_genes)
# bulkE[which(is.na(bulkE), arr.ind=T)[1]]
```

```{r, warning=FALSE}
compute_recon_error <- function(refs, verbose=F) {
  pacman::p_load('progress')
  pacman::p_load('glmnet')
  errorM <- matrix(nrow=ncol(NhoodE), ncol=length(refs))
  coefs <- matrix(nrow=ncol(NhoodE), ncol=ncol(refs[['full']])+1)

  colnames(errorM) <- names(refs)
  colnames(coefs) <- c('Intercept', colnames(refs[['full']]))

  if (!verbose)
    pb <- progress_bar$new(total = ncol(NhoodE))

  for (i in 1:ncol(NhoodE)) {
    # i <- which(colnames(NhoodE) == '40')
    if (!verbose)
      pb$tick()

    y = NhoodE[, i, drop=F]

    # fits = map(refs, function(bE) {
    for (j in 1:length(refs)) {
      # bE = refs[[1]]
      bE = refs[[j]]
      shared_genes <- intersect(rownames(y), rownames(bE))

      cvfit <- cv.glmnet(
        x=bE[shared_genes, ],
        y=y[shared_genes, ], 
        lower.limits=c(0),
        family='gaussian', alpha=.99
      )

      if (F) {
        library(e1071)
        svmfit <- svm(
          x=bE[shared_genes, ],
          y=y[shared_genes, ], 
          type="nu-regression",kernel="linear",nu=.25,
          scale=F
        )
        #get and normalize coefficients
        model <- svmfit
        dim(model$SV)
        q <- t(model$coefs) %*% model$SV
        q[which(q<0)]<-0
        w <- (q/sum(q))
        ## Compute reconstruction error 'nusvm'
        u <- sweep(bE[shared_genes, ], MARGIN=2, w, '*')
        k <- apply(u, 1, sum)
        nusvm <- sqrt((mean((k - y[shared_genes, ])^2)))
        nusvm^2
        devtools::install_github('Moonerss/CIBERSORT')
        CIBERSORT::CoreAlg
      }
        
      if (F && verbose) {
        print(i); print(j)
        print(coef(cvfit, s='lambda.1se'))
      }
      # error[i, j] <- cvfit$cvm[which.min(abs(cvfit$lambda - cvfit$lambda.1se))]
      errorM[i, j] <- 
        cvfit$cvm[which.min(abs(cvfit$lambda - cvfit$lambda.min))]

      if (F) {
        yhat <- predict(cvfit, bE)

        p <- qplot(x=y[, 1], y=yhat, alpha=.1) +
          scale_size_continuous(limits=c(.1)) +
          geom_smooth(method='rlm')
        print_plot_eval(call = { print(p) },
          filename = file.path(l_img_dir, 'test.pdf'), 
          width = 18, height = 20)
      }

      if (names(refs)[j] == 'full') {
        coefs[i, ] <- as.vector(coef(cvfit, s='lambda.1se'))
      }
    }
    if (verbose) {
      # print(errorM[i, ])
      # print(errorM[i, 'full'] / errorM[i, 'unstim_only'])
    }
  }

  return(list('errorM'=errorM, 'coefs'=coefs))
}



if (bulk_experiment == '7296') {
  all_bulk_Es <- list(
    'full'        = bulkE,
    'unstim_only' = bulkE[, grepl('medium.*medium', colnames(bulkE))],
    'only_ifny'   = bulkE[, !grepl('tgfb|mix', colnames(bulkE))],
    'only_tgfb'   = bulkE[, !grepl('ifny|mix', colnames(bulkE))]
  )
} else if (bulk_experiment == '5892') {
  all_bulk_Es <- list(
    'full'        = bulkE,
    'unstim_only' = bulkE[, grepl('unstim', colnames(bulkE))],
    'only_tnfa'   = bulkE[, grep('ifng|tgfb', colnames(bulkE), 
      invert=TRUE, value=T)],
    'only_ifny'   = bulkE[, grep('tnfa|tgfb', colnames(bulkE), 
      invert=TRUE, value=T)],
    'only_tgfb'   = bulkE[, grep('tnfa|ifng', colnames(bulkE), 
      invert=TRUE, value=T)]
  )
}

recon_error <- compute_recon_error(
  refs=all_bulk_Es, 
  # verbose=TRUE
  verbose=F
)

summary(recon_error$errorM[, 'full'] / recon_error$errorM[, 'unstim_only'])
```

```{r }
fn <- file.path(rds_dir, glue::glue('control_error_{bulk_experiment}.rds'))

if (T || !file.exists(fn)) {
  ## Select the same number of samples randomly that were selected
  ds_numbers <- setdiff(sort(map_int(all_bulk_Es, ncol)), 36L)
  sample_N <- ncol(all_bulk_Es[['full']])
  
  control_Es <- map(auto_name(ds_numbers), function(ds_N) {
    c(
      list('full' = bulkE[, setdiff(names(which(colSums(recon_error$coefs) > 0)), 'Intercept')]),
      map(1:1000, ~bulkE[, sample(1:sample_N, ds_N, replace=F)])
    )
  })

  control_error <- map(auto_name(control_Es), function(CEs) {
    compute_recon_error(refs=CEs, verbose=F)
  })
  setNames(control_error) <- as.character(ds_numbers)

  saveRDS(control_error, fn)
} else {
  control_error <- readRDS(fn)
}
```

```{r }
ord <- order(NH_test_DA$logFC)

## Virtually no difference between full model and one that's selected
## to the often selected reference samples
summary(control_error$errorM[ord, 1] / recon_error$errorM[ord, 'full'])
```

```{r}
## Minimal difference between full model and 11 randomly selected
## reference samples
n_dat <- sweep(
  x=control_error$errorM[ord, 2:1001], MARGIN=1, 
  STATS=recon_error$errorM[ord, 'full'], 
  FUN='/')
stopifnot(
  control_error$errorM[ord, 2] / recon_error$errorM[ord, 'full'] ==
  n_dat[, 1])
summary(apply(n_dat, 2, median))
```

```{r }
cn <- colnames(recon_error$coefs)

sel_mode = 'full'
sel_mode = 'high_prob'
plotM <- recon_error$coefs[ord, cn != 'Intercept' & 
  colSums(recon_error$coefs) > 0]
plotM <- sweep(plotM, MARGIN=1, FUN='/', STATS=apply(plotM, 1, sum))
if (sel_mode == 'high_prob')
  plotM <- plotM[, apply(plotM, 2, sum) > 5]

if (bulk_experiment == '7296') {
  sa <- sample_annotation_7296[match(colnames(plotM),
    sample_annotation_7296$sample_name), c('duration',
    'm_sg', 'm_sg2', 'duration2')]
} else if (bulk_experiment == '5892') {
  sa <- sa_5892[match(colnames(plotM), tolower(sa_5892$sample_name)), 
    c('duration', 'm_sg', 'm_sg2', 'duration2')]
}
sa$m_sg2 <- droplevels(sa$m_sg2)

DE <- GE_vs_stim_dur(
  M = plotM,
  min_var = 0,
  cell_width_mm = 2.5,
  scale_data = 'none',
  sa = sa,
  column_clust_method = 'none',
  row_clust_method = 'none',
  color_name = 'Profile similarity',
  plot_name = 'T_cell_associated_genes_mouse',
  show_row_dend = FALSE,
  show_column_dend = FALSE,
  show_row_names = FALSE,
  # N_hl_genes = 21,
  N_hl_genes = NULL,
  height = unit(15, 'cm'),
  show_column_names = FALSE
)

print_plot_eval(call = {
  draw(DE,
    heatmap_legend_side = 'bottom',
    annotation_legend_side='bottom')
}, filename = file.path(l_img_dir,
  glue::glue('{sel_mode}_lasso_Nhood_to_ref{bulk_experiment}.pdf')), width = 18, height = 25)
```


# Panel G


```{r }
q_ifny <- apply(cbind(
    recon_error$errorM[ord, 'only_ifny'], 
    control_error$errorM[ord, 2:1001]), 1, function(x)
  ecdf(x[2:1001])(x[1]))

q_tgfb <- apply(cbind(
    recon_error$errorM[ord, 'only_tgfb'], 
    control_error$errorM[ord, 2:1001]), 1, function(x)
  ecdf(x[2:1001])(x[1]))


make_y <- function(a = 'IFNy') {
  as.character(glue::glue('Percentile of reconstruction error with only {a} and unstim reference profiles/\nwithin distribution of reconstruction errors with identical number of randomly reference profiles'))
}

p1 <- qplot(x=NH_test_DA$logFC[ord], 
  y=q_ifny) +
  xlab('logFC DA T cell') +
  ylab(make_y(a='IFNy')) +
  geom_smooth(method='lm')
p2 <- qplot(x=NH_test_DA$logFC[ord], 
  y=q_tgfb) +
  xlab('logFC DA T cell') +
  ylab(make_y(a='TGFb')) +
  geom_smooth(method='lm') +
  theme()
print_plot_eval(call = {
  print(p1 + p2)
}, filename = file.path(l_img_dir, 'test.pdf'), 
width = 12, height = 8)
```

```{r }
ord <- order(NH_test_DA$logFC)

make_y <- function(a = 'IFNy') {
  as.character(glue::glue('Reconstruction error with only {a} and unstim profiles/\nreconstruction error with all reference profiles'))
}

# cut(q_ifny, breaks=c(0, .05, .95, 1), include.lowest=TRUE)

make_p <- function(y='only_ifny', q=q_ifny, a='IFNy') {
  q_cut <- cut(q, breaks=c(0, .025, .975, 1), include.lowest=TRUE)
  q_cut2 <- cut(q, breaks=c(0, .025, 1), include.lowest=FALSE)
  print(sum(q_cut2 == '(0, 0.025]', na.rm=TRUE))
  # print(summary(q))
  # print(summary(q_cut))
  qplot(
    x=NH_test_DA$logFC[ord], 
    y=recon_error$errorM[ord, y] / recon_error$errorM[ord, 'full'],
    alpha=.5,
    # size=1/(q + 1e-3)) +
    color=q_cut) +
  xlab('logFC DA T cell') +
  ylab(make_y(a=a)) +
  scale_alpha_continuous(guide=FALSE) +
  scale_color_manual(
    name='permutation q', 
    # breaks=c('<=0.05', 'none', '>-.95')
    labels=c('Improved', 'Same', 'Worse'),
    values=c('aquamarine3', 'grey50', 'brown3'),
    breaks=c('[0,0.025]', '(0.025,0.975]', '(0.975,1]')
  ) +
  # scale_size_continuous(
  #   name='log10 permutation q', 
  #   labels=function(x) round(-log10(x), 3),
  #   range=c(0.5, 2)
  # ) +
  # geom_smooth(method='lm', show.legend=F)
  theme()
}
p1 <- make_p(y='only_ifny', q=q_ifny, a='IFNy')
p2 <- make_p(y='only_tgfb', q=q_tgfb, a='TGFb')

print_plot_eval(call = {
  print(p1 + p2)
}, filename = file.path(l_img_dir, 'test.pdf'), 
width = 17.4, height = 10)
```

```{r, warning=FALSE}
t.test(x = error[ord, 1], y = error[ord, 2])
t.test(x = error[ord, 1], y = error[ord, 3], paired = TRUE)

summary(error[ord, 2] / error[ord, 1])
summary(error[ord, 3] / error[ord, 2])
summary(error[ord, 3] / error[ord, 1])
```
