---
title: "Gene set based time inference"
author: "Maarten Slagter"
date: "`r Sys.time()`"
output:
  bookdown::pdf_document2:
    fig_caption: yes
    toc: false
params:
  experiment: 6369
  so: NA
  GS_p_dat: NA
  cn_mode: NA
  report_cp: NA
  p_values: NA
  all_univariate_gs_comps: NA
---

This Rmarkdown makes plots pertaining to time inference based on humble gene
set scores. Set the idx variable to 8 for Figure 1F(experiment 6369); 9(?) for
Figure 1G (experiment 6489); 17 for Figure 3D (experiment 6493); 

```{r, warning=FALSE}
options('plot_force_file' = TRUE)
 
if (interactive()) {

  source('~/MirjamHoekstra/R/init.R')
  if (!exists('idx')) {
    idx = 4 #5310 in vitro
    idx = 8 #6369_simple_reference
    idx = 16 #6601 simple
    idx = 17 #6493 simple reference
  }
  source(file.path(r_dir, 'get_knit_params.R'))
  tab <- targets_env$gs_param_table
  load_prereqs(param_grid = tab, idx = idx)
} else {
  # attach(params)
  experiment <- params$experiment
  so <- params$so
  GS_p_dat <- params$GS_p_dat
  cn_mode <- params$cn_mode
  report_cp <- params$report_cp
  p_values <- params$p_values
  all_univariate_gs_comps <- params$all_univariate_gs_comps
}

source(file.path(r_dir, 'GS_time_inference.R'))
eval(time_inference_preproc)
```

# UMAP

```{r umaps, warning=FALSE}
p1 <- DimPlot(
  so,
  # group.by = group_var_cn_mode[[cn_mode]],
  group.by = 'stim_group',
  label.size = 2,
  pt.size = 8, label = FALSE,
  # raster.dpi = c(512, 512).
  raster.dpi = 4 * c(512, 512),
  split.by = group_vars[length(group_vars)],
  # split.by = 'stim_group',
  raster = TRUE) +
  # scale_colour_stim_group(so@meta.data) +
  theme_cyto_inf() +
  ggtitle('') +
  guides(color = guide_legend(ncol = 2))
print_plot_eval(
  plot(p1),
  width = length(unique(so@meta.data$duration)) * 17.4 / 3,
  height = 10,
  filename = file.path(out_dir,
    glue::glue('exp{report_cp}_UMAP.pdf')))
```

```{r umaps-stim-group, warning=FALSE}
so_o <- tar_read_raw(glue::glue('filtered_cleaned_so_{experiment}'))

p1 <- DimPlot(
  so_o,
  # group.by = group_var_cn_mode[[cn_mode]],
  group.by = 'stim_group',
  label.size = 2,
  pt.size = 8, label = FALSE,
  # raster.dpi = c(512, 512).
  raster.dpi = 4 * c(512, 512),
  split.by = group_vars[length(group_vars)],
  # split.by = 'stim_group',
  raster = TRUE) +
  # scale_colour_stim_group(so@meta.data) +
  theme_cyto_inf() +
  ggtitle('') +
  guides(color = guide_legend(ncol = 2))
print_plot_eval(
  plot(p1),
  width = length(unique(so@meta.data$duration)) * 17.4 / 3,
  height = 10,
  filename = file.path(out_dir,
    glue::glue('exp{report_cp}_var_genes_UMAP.pdf')))
```


# Expression heatmaps

```{r expression-heatmaps}
fn <- file.path(out_dir,
    glue::glue('expression_heatmaps-{report_cp}.pdf'))
mean_expression_heatmaps(GS_p_dat, o_fn = fn)
```

```{r expression-heatmaps-cn}
exclude_gs <- 'HK_genes'
exclude_gs <- c()
if (length(remaining_group_vars) > 0) {
  p <-
    GS_p_dat %>%
    tidyr::pivot_longer(cols = 1:35, names_to = 'gs') %>%
    dplyr::group_by_at(c('condition_name', 'gs', remaining_group_vars)) %>%
    dplyr::summarize(
      feat = median(value),
      perc_exp = mean(value > 0)
    ) %>%
    # dplyr::group_by(gs) %>%
    # dplyr::mutate(feat = (feat - median(feat)) / IQR(feat)) %>%
    dplyr::filter(!is.na(feat) & is.finite(feat)) %>%
    dplyr::filter(!gs %in% exclude_gs) %>%
    ggplot(aes(x = condition_name, y = gs,
        fill = feat,
        label = round(feat, 2),
        )) +
      rotate_x_labels(45) +
      geom_tile() +
      geom_text(size = 2.2) +
      # scale_fill_viridis_c(direction = -1L) +
      scale_x_discrete(name = 'Condition', expand = c(0, 0)) +
      scale_y_discrete(name = 'Gene set', expand = c(0, 0)) +
      facet_wrap(remaining_group_vars) +
      guides(fill = 'none')

  fn <- file.path(out_dir,
    glue::glue('expression_heatmaps-{report_cp}-all_in_one.pdf'))
  print_plot_eval(print(p),
    width = 17.4, height = 20, filename = fn)
}
```

# Violins

Unfiltered vs. bottom 90 (with respect to a GS of interest, either
'general' IFNy or TNFa) or top 10

```{r violins, warning=FALSE}
if (interactive() && !test_rendering())
  source('~/MirjamHoekstra/R/init.R')

# GS_p_dat %>%
#   dplyr::distinct(stim_group, duration, Ag, grp_i)

  tidyr::expand_grid(
    # perc = c(.75, .9),
    # perc = c(.95),
    # q_gs = all_fns[1:17],
    # q_gs = all_fns[1],
    # q_gs = paste0(c('IFNy.plateau', 'IFNy.late', 'TNFa.early',
    #     'TNFa.6h.max', 'TNFa.late'), '.vanilla.unweighted.sum')
    # q_gs = c('IFNy.plateau', 'IFNy.late', 'TNFa.early',
    #     'TNFa.late', 'IFNy', 'TNFa', 'synergy')
    q_gs = c('IFNy', 'TNFa', 'synergy')
  ) %>%
  # tidyr::expand_grid(
  #   perc = NA_real_,
  #   anchor_gs = NA_character_,
  #   stim_group_constant = NA_character_,
  #   duration_constant = NA_character_
  # ) %>%
  tidyr::expand_grid(
    perc = c(.90),
    # anchor_gs = c('IFNy', 'TNFa')
    anchor_gs = c('IFNy')
  ) %>%
  {
    if (F && length(setdiff(GS_p_dat$duration, c('Unknown', NA))) > 1) {
      tidyr::expand_grid(
        .,
        duration_constant = c(levels(GS_p_dat$duration),
          NA_character_),
        stim_group_constant = c(levels(GS_p_dat$stim_group),
          NA_character_),
      )
   } else {
      tibble(., stim_group_constant = NA_character_,
        duration_constant = NA_character_)
    }
  } %>%
  # dplyr::filter(
  #   is.na(stim_group_constant) | is.na(duration_constant)) %>%
  dplyr::mutate(i = 1:n()) %>%
  purrr::pmap(function(perc, q_gs, anchor_gs,
      stim_group_constant, duration_constant, i) {
    if (!q_gs %in% colnames(GS_p_dat)) {
      rlang::warn(glue::glue('{q_gs} not found'))
      return(NULL)
    }

    xl <- scale_y_continuous(limits = range(GS_p_dat[[q_gs]]))

    if (i == 1) {
      ## Make legend
      mk_ann_GS_vln(
        p_dat = GS_p_dat,
        cn_mode = cn_mode,
        fn = q_gs,
        out_dir = out_dir,
        experiment = experiment,
        return_mode = 'file',
        p_values = p_values,
        plot_lgd = TRUE
      )
    }

    format_gs <- function(x) stringr::str_replace_all(x, '/| ', '_')
    anchor_gs_p <- format_gs(anchor_gs)
    q_gs_p <- format_gs(q_gs)

    l_dat <- GS_p_dat
    p_y_loc <- 
      quantile(l_dat[[as.character(q_gs)]], .9999, na.rm = T) * 1.00

    # table(l_dat$Ag)
    if (!is.na(stim_group_constant)) {
      s_ <- tibble(stim_group = factor(stim_group_constant,
          levels = levels(l_dat$stim_group)))
      l_dat <- l_dat %>%
        dplyr::right_join(s_, by = 'stim_group')
    }
    if (!is.na(duration_constant)) {
      s_ <- tibble(duration = factor(duration_constant,
          levels = levels(l_dat$duration)))
      l_dat <- l_dat %>%
        dplyr::right_join(s_, by = 'duration')
      # table(l_dat$Ag)
    }

    gvars <- colnames(l_dat) %>%
      intersect(c('stim_group', 'duration', 'Ag', 'frozen'))
    N_conditions <- l_dat %>%
      group_by_at(gvars) %>%
      dplyr::summarize(n()) %>%
      nrow()

    p0 <-
      mk_ann_GS_vln(
        p_dat = l_dat,
        cn_mode = cn_mode,
        fn = q_gs,
        experiment = experiment,
        p_y_loc = p_y_loc,
        return_mode = 'p',
        p_values = p_values,
        plot_lgd = FALSE
      )
    p0[[attr(p0, 'N_legends')+1]] <-
      p0[[attr(p0, 'N_legends')+1]] + xl +
      ggtitle(q_gs)

    g1 <- retrieve_anchor_thresh(
      dtf = GS_p_dat,
      ref_condition = ref_condition,
      perc = perc,
      anchor_gs = anchor_gs
    )
    for (cn in colnames(ref_condition)) {
      g1[[cn]] <- NULL
    }

    if ('duration' %in% colnames(g1) &&
        is.factor(GS_p_dat$duration)) {
      g1$duration <- factor(g1$duration,
        levels = levels(GS_p_dat$duration))
    }
    # for (cn in intersect(colnames(g1), colnames(GS_p_dat))) {
    #   as(g1[[cn]], class(GS_p_dat[[cn]]))
    # }
    if (all(is.na(g1$duration))) {
      ofn <- file.path(out_dir,
        glue::glue('filtered_violins-{report_cp}\\
          {make_flag(q_gs_p)}\\
          .pdf')
        )
      print_plot_eval(
        print(p0),
        width = 17.4,
        # height = ifelse(experiment %in% c('6369', '6489'), 50, 25),
        height = N_conditions + 6,
        filename = ofn
      )
    } else {
      p1 <-
        l_dat %>%
        dplyr::inner_join(
          g1,
          by = intersect(colnames(g1), colnames(l_dat))
          ) %>%
        dplyr::filter(.data[[anchor_gs]] < q9) %>%
        mk_ann_GS_vln(
          cn_mode = cn_mode,
          fn = q_gs,
          experiment = experiment,
          p_y_loc = p_y_loc,
          return_mode = 'p',
          p_values = p_values,
          plot_lgd = FALSE
        ) +
        plot_annotation(caption = '< than q9 of {anchor_gs_p}')
      p1[[attr(p1, 'N_legends')+1]] <-
        p1[[attr(p1, 'N_legends')+1]] + xl +
        ggtitle(glue::glue('Lower than or equal to {perc*100}th \\
            percentile of {anchor_gs} of unstim control'))

      p2 <-
        l_dat %>%
        dplyr::inner_join(
          g1,
          by = intersect(colnames(g1), colnames(l_dat))
          ) %>%
        dplyr::filter(.data[[anchor_gs]] >= q9) %>%
        mk_ann_GS_vln(
          cn_mode = cn_mode,
          fn = q_gs,
          experiment = experiment,
          p_y_loc = p_y_loc,
          return_mode = 'p',
          p_values = p_values,
          plot_lgd = FALSE
          ) +
        plot_annotation(caption = '>= than q9 of {anchor_gs_p}')
      p2[[attr(p2, 'N_legends')+1]] <-
        p2[[attr(p2, 'N_legends')+1]] + xl +
        # ggtitle(glue::glue('Top {scales::percent(1-perc)} w.r.t {anchor_gs}'))
        ggtitle(glue::glue('Higher than {perc*100}th percentile of \\
            {anchor_gs} of unstim control'))

      stim_group_constant <- format_gs(stim_group_constant)
      duration_constant <- format_gs(duration_constant)
      ofn <- file.path(out_dir,
        glue::glue('filtered_violins-{report_cp}\\
          {make_flag(q_gs_p)}\\
          {make_flag(stim_group_constant)}\\
          {make_flag(duration_constant)}\\
          {make_flag(anchor_gs_p)}\\
          {make_flag(perc)}\\
          .pdf')
        )

      print_plot_eval(
        print(p0 / p1 / p2),
        width = 17.4,
        # height = ifelse(experiment %in% c('6369', '6489'), 50, 25),
        height = N_conditions*3 + 6,
        filename = ofn
      )
    }
  })
```

# AUCs of distinguishing capabilities of gene sets

All vs. all conditions

```{r auroc-all-conditions-prep}
library(tidyr)
source('~/MirjamHoekstra/R/init.R')

gst2gs <-
  list(
    'default' = names(tar_read(default_gene_sets)),
    'time_informative' = names(tar_read(mirjam_genes)),
    # 'time_informative' = names(tar_read(time_informative_genesets)),
    'clustered_mono' = names(tar_read(clustered_mono_genes)),
    'clustered_IFNy' = names(tar_read(clustered_IFNy_genes)),
    'clustered_TNFa' = names(tar_read(clustered_TNFa_genes)),
    'clustered_limma_MR' = names(tar_read(clustered_limma_MR_genes))
  ) %>%
  tibble::enframe('type', 'gs') %>%
  tidyr::unnest(cols = gs)

univariate_gs_comps <-
  all_univariate_gs_comps %>%
  # head() %>%
  dplyr::right_join(gst2gs, by = 'gs') %>%
  tidyr::unnest_wider(comp_levs) %>%
  dplyr::mutate(
    cn1 = map_chr(condition_name, ~.x[[1]]),
    cn2 = map_chr(condition_name, ~.x[[2]])
  ) %>%
  dplyr::select(-condition_name, -comp_name, -type) %>%
  dplyr::filter(!is.na(AUROC)) %>%
  { . }
```

```{r auroc-all-conditions, warning=FALSE}
# dplyr::rename(univariate_gs_comps, cn1 = cn2, cn2 = cn1) %>%
#   head()
expected_conditions <- 
  unique(as.character(GS_p_dat[[group_var_cn_mode[[cn_mode]]]]))
univariate_gs_comps
```

# Time scatters

```{r time-scatters-norm}
if (interactive() && !test_rendering())
  source('~/MirjamHoekstra/R/init.R')

if (experiment %in% c('6489', '6369', '6493')) {
  ## First make plot of unnormalized ref
  subs_conditions <- list(ref_condition, subs_condition,
    subs_condition, subs_condition)
  ref_conditions <- list(NULL, ref_condition, ref_condition,
        ref_condition)
  tibble(
    description = c('unstim_ref', 'all_cells', 'top10', 'top10'),
    perc = c(NA_real_, NA_real_, .90, .90),
    anchor_gs = c(NA_character_, NA_character_, 'IFNy', 'TNFa')
  ) %>%
  dplyr::mutate(i = 1:n()) %>%
  purrr::pmap(function(description, perc, anchor_gs, i) {
    p_dat <-
      GS_p_dat %>%
      nest_by(duration) %>%
      dplyr::summarize(
        data %>%
        normalize_gs(
          ref_condition = ref_conditions[[i]]
        ) %>%
        filter_anchor_thresh(
          anchor_gs = anchor_gs,
          mode = '>=',
          ref_condition = ref_conditions[[i]],
          perc = perc
        ) %>%
        { . }
      ) %>%
      { . }

    plots <-
      p_dat %>% {
        if (!is.null(subs_conditions[[i]]))
          dplyr::right_join(., subs_conditions[[i]])
        else .
      } %>%
      make_all_tgs_scatter_plots(
        colour_var = 'duration',
        balanced_axes = F,
        comps = tribble(
          ~x_var, ~y_var,
          'IFNy', 'TNFa',
          'IFNy.plateau', 'IFNy.late',
          'IFNy.late', 'IFNy.plateau',
          'TNFa.early', 'TNFa.late'
        )
      )

    fn <- file.path(out_dir,
      glue::glue('time_gs_scatters\\
        {make_flag(experiment)}\\
        {make_flag(description)}\\
        {make_flag(anchor_gs)}\\
        {make_flag(perc)}.pdf')
      )
    plot_panel_layout(
      plots = plots, ncol = 2, nrow = 3, clear_redundant_legends = F,
      width = 17.4, height = 25, filename = fn, labels = NULL
    )
  })
}
# print_plot_eval(
#   print(patchwork::wrap_plots(
#     plots, ncol = 2, nrow = 3, guides = 'collect'
#   )),
#   width = 17.4, height = 25,
#   filename = fn)
```

Predictive capacity of gene set scores

```{r, eval=F}
b_so <- tar_read(bulk_5029_so)
b_so_s <- ComputeGeneSetScores(
  so = b_so,
  weighting_funcs = all_weighting_funcs['unweighted'],
  genesets = tar_read(all_genesets),
  simplify_names = F
)
b_dat <- assay2dtf(b_so_s, assay = 'GS')
colnames(b_dat) <-
  stringr::str_replace(colnames(b_dat),
    '.vanilla.unweighted.sum', '')

s_dat <- tar_read(GS_p_dat_5310_simple_reference)
colnames(s_dat) <-
  stringr::str_replace(colnames(s_dat),
    '.vanilla.unweighted.sum', '')

e_vars <- c('TNFa', 'IFNy', 'synergy', 'IFNy.late', 'IFNy.plateau', 'TNFa.early', 'TNFa.late')
# cor(t(b_dat[, e_vars]), t(s_dat[, e_vars]))

# M <- cbind(t(b_dat[, e_vars]), t(s_dat[, e_vars])) %>%
#   as.matrix()
# class(t(s_dat[, e_vars]))
# outer(t(t(b_dat[, e_vars])), t(t(s_dat[, e_vars])), cosine)
# cosine(M)

CM <- matrix_cosine(x = t(b_dat[, e_vars]), y = t(s_dat[, e_vars]))

gamma <- 1e-5
gamma <- 2e0
CMe <- exp(-gamma * CM)
CM_NN <- apply(CMe, 2, function(x) ifelse(x == max(x), 1, 0))
stopifnot(apply(CM_NN, 2, sum) >= 1)

reg_vars <- c('duration', 'ifn_conc', 'tnf_conc')
ref_sa <- extract_sa(b_so_s, 
  meta_fields = reg_vars) %>%
  numerify_regressors() %>%
  norm_regressors()

query_sa <- s_dat %>%
  dplyr::select(any_of(reg_vars)) %>%
  numerify_regressors() %>%
  norm_regressors()

error_Ms <- map(auto_name(reg_vars), function(rv) {
  outer(ref_sa[[rv]], query_sa[[rv]], '-')
})

by_var <- c('condition_name')
query_rsa <- s_dat %>%
  dplyr::select(any_of(by_vars), duration, stim_group) %>%
  order_condition_name()
query_rsa[[by_var]]

WM <- CM_NN 
parm_grid <- tidyr::expand_grid(
  cn = levels(query_rsa[[by_var]]),
  rv = reg_vars
) 
# sum_funcs <- list(median)
parm_grid$error <- parm_grid %>%
  purrr::pmap(function(cn, rv) {
    ## Select cells corresponding to condition
    idxs <- which(query_rsa[[by_var]] == cn)
    if (length(idxs) == 0) return(NULL)
    ## Slice error matrix and multiply by weight matrix
    {
      error_Ms[[rv]][, idxs] * WM[, idxs]
    } %>%
    apply(2, sum) %>%
    # map_dfc(sum_funcs, ~.x(.))
    { . }
  })
parm_grid <- parm_grid %>%
  dplyr::filter(!sapply(error, is.null))
parm_grid$error_median <- map_dbl(parm_grid$error, function(x) {
  if (is.null(x)) return(NA_real_)
  tryCatch(median(abs(x), na.rm = T), error = function(e) { NA_real_ }) 
})
parm_grid <- parm_grid %>%
  dplyr::filter(!sapply(error_median, is.na))
parm_grid$error_q1 <- map_dbl(parm_grid$error, function(x) {
  if (is.null(x)) return(NA_real_)
  tryCatch(quantile(abs(x), .1, na.rm = T), error = function(e) { NA_real_ }) 
})
parm_grid$error_q9 <- map_dbl(parm_grid$error, function(x) {
  if (is.null(x)) return(NA_real_)
  tryCatch(quantile(abs(x), .9, na.rm = T), error = function(e) { NA_real_ }) 
})
```
